---
title: "Plotting Milk Tea, GV-Cinema and Macdonalds locations"
author: "JD"
output:
  html_document:
    keep_md: yes
    theme: cerulean
    code_folding: show
    toc: true
    toc_depth: 4
---

<style type="text/css">
.main-container{
  max-width: 1300px;
  margin-left: auto;
  margin-right: auto;
}
body, td{
  font-family: Helvetica;
  font-size: 16pt;
}
pre{ 
  font-size: 14px; 
}
pre.r{ 
  font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=15)
work_directory = getwd()

# specify the packages of interest
packages = c("tidyverse","readxl","geosphere","rvest","pander","knitr","jsonlite","kableExtra")


# use this function to check if each package is on the local machine
# if a package is installed, it will be loaded
# if any are not, the missing package(s) will be installed and loaded
package.check <- lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
        install.packages(x, dependencies = TRUE, repos = 'http://cran.us.r-project.org')
        library(x, character.only = TRUE)
    }
})

# installs ggmap directly from github source. NO NOT use cran version
if(!requireNamespace("devtools")) install.packages("devtools")
devtools::install_github("dkahle/ggmap", ref = "tidyup")
library(ggmap)

search()
```

<br/>

**References**  
*put all reference links here*

***
<br />

## Getting the Data 

We first create our index which holds the relevant locations and the urls we plan on scraping.

```{r}
# reading required variables 
# either have a valid google api key for geocoding api or have "lat_long_data.csv" in your working directory
creds = read_json(paste(path.expand('~'),'/creds.json', sep = ''))
google_apikey = creds$Google_api$geocode_api
register_google(key = google_apikey)

rm(creds)


# setting the reference index
locations <- c("Liho", "Koi", "GV Cinema", "McDonalds", "Gong cha")
urls <- c("https://www.streetdirectory.com/businessfinder/company_branch/163304/5890/",
          "https://www.streetdirectory.com/businessfinder/company_branch/109085/168981/",
          "https://www.streetdirectory.com/businessfinder/company_branch/36232/30314/",
          "https://www.streetdirectory.com/businessfinder/company_branch/65869/254129/",
          "https://www.gong-cha-sg.com/stores/")
          
index <- data.frame(locations, urls, stringsAsFactors = FALSE) 
kable(index) %>% kable_styling()
```

<br />

We then define the relevant functions we plan on using to get the data needed. Note the created function already alters the output into the format that we want. 

```{r}
# function used to scrape data 
getting_data <- function(url_link, css_selector) {
  url_link %>%
    read_html() %>%
    html_nodes(css_selector) %>%
    html_text() %>%
    str_replace_all(.,"\\r|\\n","") %>%
    str_trim() %>%
    str_replace_all(.,"  ","")
}


# reading data for Liho, Koi, GV Cinema and McDonalds
basedata <-
  map(seq_along(index$urls[1:4]), function(x) {
    data.frame(name = getting_data(index$urls[x], ".company_branch_name"),
               address = getting_data(index$urls[x], ".company_branch_address"),
               postcode = str_extract(getting_data(index$urls[x], ".company_branch_address"), "[0-9]{6}$"),
               indicator = index$locations[x], 
               stringsAsFactors = FALSE)
    }) %>%
  setNames(index$locations[1:4])


# reading Gong cha data
# this is different as the Gong cha website loads using javascript
# I found no elegant way to read the data directly from R so defaulted to the easiest method
if(!file.exists("gong_cha.html")) {
warning("You need to save http://www.gong-cha-sg.com/stores/ into working directory as 'gong_cha.html'")
  
  } else {
    gong_cha_data <-
      data.frame(name = getting_data("gong_cha.html", ".p-title"),
                 address = getting_data("gong_cha.html", ".p-area"),
                 postcode = str_extract(getting_data("gong_cha.html", ".p-area"), "[0-9]{6}$"),
                 indicator = index$locations[5],
                 stringsAsFactors = FALSE)
    
    basedata$`Gong cha` <- gong_cha_data
}

location_data <-
  bind_rows(basedata)


# check if there are any rows with postalcode not equal to 6 digits
location_data %>% mutate(count_postal = nchar(postcode)) %>% filter(count_postal != 6)


# table output 
kable(location_data %>% group_by(indicator) %>% slice(1:3)) %>% kable_styling()
```

<br />

We next use the google maps geocode api [Google maps gecoding intro](https://developers.google.com/maps/documentation/geocoding/intro) to get the relevant latitudes and longitudes. To do this you need to have a valid api key. If you do not have one, fret not. I have already pre-saved the geocode information before hand. Just make sure you have the reference csv "lat_long_data_with_errors" in your working enviroment. 


```{r}
# function used to call the api
calling_google_geocode <- function(postcode, apikey) {

  output <-
    "https://maps.googleapis.com/maps/api/geocode/json?address=Singapore+xxpostcodexx&key=xxapikeyxx" %>%
    str_replace_all(c('xxpostcodexx' = postcode, 'xxapikeyxx' = apikey)) %>%
    fromJSON()
  
  
  if(length(output) == 0) return(data.frame())
  return(output$results$geometry$location)
}


# getting latitudes and longitudes
if(!file.exists("lat_long_data_with_errors.csv")) {
  
  lat_long_data <-
  location_data %>%
  .$postcode %>%
  map(., function(i) {
    output_api <- calling_google_geocode(postcode = i, apikey = google_apikey)
    return(data.frame(postcode_used = i, output_api, stringsAsFactors = FALSE))
    
  }) %>%
  bind_rows() %>%
  bind_cols(location_data, .)
  
  # saving data cause I am cheap and google apis can actually charge you now
  lat_long_data %>% write_csv("lat_long_data_with_errors.csv")
  
  } else {
  
  lat_long_data <- read_csv("lat_long_data_with_errors.csv")
  
  }


# table output 
kable(lat_long_data %>% select(-postcode_used) %>% group_by(indicator) %>% slice(1:3)) %>% kable_styling()
```

***
<br />

## Visualising the Data 

Now we get a static map of singapore and then plot the latitude and longitude points onto the map. Unfortunately an error has occured. Can you spot it ?

```{r, fig.width = 14, fig.height = 9, out.width = "2000", out.height = "500"}

if(!file.exists("singapore_map.rds")) {
  
  map_singapore <- 
    get_map(location = 'Singapore', zoom = 12, color = 'bw')
  
  saveRDS(map_singapore, file = "singapore_map.rds")
  } else {
  
  map_singapore <- readRDS("singapore_map.rds")
  }

# plotting 
ggmap(map_singapore, extent = "device") +
  geom_point(data = lat_long_data, aes(x = lng, y = lat, colour = indicator), size = 2, position = position_jitter(w = 0.002, h = 0.002)) +
  scale_color_manual(breaks = c("Gong cha", "GV Cinema", "Koi", "Liho", "McDonalds"),
                        values=c("darkred", "darkgreen", "darkblue", "darkgoldenrod4", "pink2")) +
  geom_point(aes(x = 103.844185, y = 1.387563), shape = 18, size = 4, colour = 'red') +
  annotate("text", x = 103.844185, y = 1.39, label = 'bold("NCS Hub")', size = 3.5, parse = TRUE)
```

<br />

For some reason, google's geocode api cannot get the latitude and longitude values for these postcodes and defaults to the value for the city of Singapore, which happens to be in MacRitchie <span style = "color:Maroon">***(that cluster of points in the middle of the map)***</span>. Below are the values that were geocoded wrongly. You can try imputting the postcodes into google maps directly and see if you get anything.

```{r}
# filtering out the errors from the google api
lat_long_data %>% 
  filter(lat == 1.3520830,lng == 103.8198360) %>% 
  kable() %>%
  kable_styling() %>%
  kableExtra::scroll_box(width = "100%", height = "400px")
  
```

***
<br />

## Correcting Errors

Thankfully, the google api is clever enough to guess the latitude and longitude values from the address string. Unfortunately, the api call function we created doesnt accomodate such strings. No worry, `geocode()` comes to the rescue, automatically formats address strings for the google api.   


```{r}

wrong_data <-
  lat_long_data %>% 
  filter(lat == 1.3520830,lng == 103.8198360) %>%
  select(-lat,-lng)


if(!file.exists("correcting_error.rds")) {
  correcting_error <-
    wrong_data %>%
    .$address %>%
    map(function(i) {
      
      gcd = geocode(i, output = "latlon", source = "google")
      }) %>%
    bind_rows() %>%
    select(lat, 'lng' = lon) %>%
    bind_cols(wrong_data, .)
  
  saveRDS(correcting_error, file = "correcting_error.rds")
  } else {
  
  correcting_error <- readRDS("correcting_error.rds")
  
  }

kable(correcting_error) %>%
  kable_styling() %>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```

***
<br />

## Replotting and Distance analysis

We then look into which tea shop, cinema or macdonalds is closest to the office. We used the haversine formula from the geosphere package to estimate direct line distance which offers a decent comparative guage. 

```{r echo=TRUE}

ncs_hub <-
  data.frame(lat = 1.387563, lng = 103.844185)


distance_data <-
  lat_long_data %>% 
  filter(lat != 1.3520830,lng != 103.8198360) %>%
  bind_rows(., correcting_error) %>%
  rowwise() %>%
  mutate(distance = round(distHaversine(c(lng,lat), c(ncs_hub$lng,ncs_hub$lat)),2))

write_csv(distance_data, "lat_long_data_correct.csv")

```

<br />

A quick replot of the corrected data shows the errors have been fixed.  

```{r, fig.width = 14, fig.height = 9, out.width = "2000", out.height = "500"}
ggmap(map_singapore, extent = "device") +
  geom_point(data = distance_data, aes(x = lng, y = lat, colour = indicator), size = 2, position = position_jitter(w = 0.002, h = 0.002)) +
  scale_color_manual(breaks = c("Gong cha", "GV Cinema", "Koi", "Liho", "McDonalds"),
                        values=c("darkred", "darkgreen", "darkblue", "darkgoldenrod4", "pink2")) +
  geom_point(aes(x = 103.844185, y = 1.387563), shape = 18, size = 4, colour = 'red') +
  annotate("text", x = 103.844185, y = 1.39, label = 'bold("NCS Hub")', size = 3.5, parse = TRUE)
```

<br />

We then look at which top 3 tea shop, cinema and macdonalds are closest to the office. Naturally the results are not surprising. But its good to have verification.

```{r}
distance_data %>%
  group_by(indicator) %>% 
  arrange(distance) %>% 
  slice(1:3) %>% 
  kable() %>%
  kable_styling()
```


***
<br />

## Conclusion

Geospatial stuff is fun. This is merely the beginning. If you liked my work go to my [github page] https://github.com/jaredgoh/Project-Work for more guides. Its a young repo, but it will grow !

***
<br />

